Databrary: Enabling sharing and reuse of research video

### 2015-04-16 jcdl-15/demo.tex
minor edits, think it's ready after a few last passes

Databrary began accepting contributions in early 2014 and opened for general use in October 2014.
As of April 2015, it hosts 7,340 video files totaling 2,409 hours of recordings along with 2,647 additional files.
These files make up 3,108 study sessions and are covered by 2,000 metadata records (including 1,804 individual participants).
Data originate from 35 individual contributors across 25 different universities.

@@

Focusing on a particular research domain allows us to develop tools informed by how the intended community of users currently works, while accommodating the specific privacy concerns for data that often has a high disclosure risk.
We have created of a set of data management features that empower researchers to actively manage their own projects---to upload data with accompanying metadata---as each study unfolds. 
These features include a spreadsheet interface for entering, editing, and viewing session-level metadata: participant demographics, conditions of study, tasks in the experiment, permission levels, study groups and conditions, etc. 
Most researchers use desktop spreadsheets for precisely this purpose in their own labs, making the interface and functionality intuitive to users. 
We have also implemented a timeline for uploading, viewing, and tagging video assets related to sessions.
The timeline view is designed to look and operate like video editing software commonly used in many research labs.
It allows users to upload video files, position them to reflect the temporal order of each component of data collection, and annotate video sections with user-generated tags.

@@

In addition, because video data involving human subjects is identifiable, a core challenge to storing and sharing this type of research data is creating the access checks necessary to honor the participant consents under which the data were collected. 
User registration for gaining access to Databrary involves an authorization process by which site users are verified to be faculty, students, or researchers affiliated with an educational institution.
Users are required to review and sign an access agreement stating that they will treat the data they are able to access through Databrary in a manner that complies with the ethical handling of research data involving human subjects\footnote{http://databrary.org/access/policies/agreement.html}.
Finally, contributors are offered permission levels for files and groups of files that allow them to fine-tune who has access to their data.

\section{Discovery, Reuse, and Beyond}

Databrary's mission is not merely the storage and management of research data, but also to make these data discoverable by other researchers and to facilitate their reuse.
Currently we support text searching over the entire repository, and we plan to continue implement more refined searches that allows faceting of search results based on tags, keywords, and session metadata such as participant age, task, study condition, and context.
Additionally, Databrary allows for the annotation of videos and video segments with tags, which can serve as an index point for more refined search functionality.

We aim to allow data to be reused in future studies, cited in scholarly communication, as well as used an aid in scientific and educational presentations.
To aid in using datasets as a bibliographic source, each dataset comes with a formatted citation, a persistent URI, and soon will also be assigned a Digital Object Identifier (DOI) for referencing in scholarly publications and presentations.
Based on input from site users we have also built in the ability to excerpt or highlight clips from videos in such a way that they might serve as an exemplar clip for the dataset or be used in the classroom as an education tool or at professional conferences.

As we continue to gather feedback from our intended user base, we not only see Databrary as a tool that facilitates the standardized management of research data, but also as an opportunity to enhance the ability for academics communicate about and augment their research.
We have already addressed the basic upload, data management, and storage needs of our target community, and have sufficiently answered the question of how to protect sensitive video data while providing wider opportunities to share those data.
Going forward, we hope to focus more on developing the tools for searching, discovering, and reusing those data, allowing researchers to work with their data in ways they had not previously imagined.

>>>>

Databrary began accepting contributions in early 2014 and opened for general use in October 2014.
As of April 2015, it hosts 7,300 video files totaling 2,400 hours of recordings along with 2,600 additional files.
These files make up 3,100 study sessions and are covered by 2,000 metadata records (including 1,800 individual participants).
Data originate from 35 individual contributors across 25 different universities.

@@

Focusing on a particular research domain allows us to develop tools informed by how the intended community of users currently works, while accommodating the specific privacy concerns for data that often has a high disclosure risk.
We have created a set of data management features that empower researchers to actively manage their own projects---to upload data with accompanying metadata---as each study unfolds. 
These features include a spreadsheet interface for entering, editing, and viewing session-level metadata: participant demographics, conditions of study, tasks in the experiment, permission levels, study groups and conditions, etc. 
Most researchers use desktop spreadsheets for precisely this purpose in their own labs, making the interface and functionality intuitive to users. 
We have also implemented a timeline for uploading, viewing, and tagging video assets related to sessions.
The timeline view is designed to look and operate like video editing software commonly used in many research labs.
It allows users to upload video files, position them to reflect the temporal order of each component of data collection, and annotate video sections with user-generated tags.

@@

In addition, because video data involving human subjects is identifiable, we have built the access checks necessary to honor the participant consents under which the data were collected. 
User registration for gaining access to Databrary involves an authorization process by which site users are verified to be faculty, students, or researchers affiliated with an educational institution.
Users are required to review and sign an access agreement stating that they will treat the data they are able to access through Databrary in a manner that complies with the ethical handling of research data involving human subjects\footnote{http://databrary.org/access/policies/agreement.html}.
Finally, contributors are offered permission levels for files and groups of files that allow them to fine-tune who has access to their data.

\section{Discovery and Reuse}

Databrary's mission is not merely the storage and management of research data, but also to make these data discoverable by other researchers and to facilitate their reuse.
Currently we support text searching over the entire repository, and we plan to continue to implement more refined searches that allows faceting of search results based on session metadata such as participant details, tasks involved in the experiment, study conditions, and context.
Additionally, Databrary allows for the annotation of videos and video segments with tags, which can serve as an index point for more refined search functionality.

We aim to allow data to be reused in future studies, cited in scholarly communication, as well as used to aid in scientific and educational presentations.
To promote using datasets as a bibliographic source, each dataset comes with a formatted citation, a persistent URI, and soon will also be assigned a Digital Object Identifier (DOI) for referencing in scholarly publications and presentations.
Based on input from site users we have also built in the ability to excerpt or highlight clips from videos in such a way that they might serve as an exemplar clip for the dataset or be used in the classroom as an education tool or at professional conferences.

As we continue to gather feedback from our intended user base, we not only see Databrary as a tool that facilitates the standardized management of research data, but also as an opportunity to enhance the ability for academics to communicate about and augment their research.
We have already addressed the basic upload, data management, and storage needs of our target community, and have sufficiently answered the question of how to protect sensitive video data while providing wider opportunities to share those data.
Going forward, we hope to focus more on developing the tools for searching, discovering, and reusing those data, allowing researchers to work with their data in ways they had not previously imagined.
